---
title: "Assigning the distance to MPA boundaries to all GFW points of fishing activity"
output: html_notebook
---

```{r include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warnings = FALSE, message = FALSE)

# Necessary libraries
library(tidyverse)
library(bigrquery)
library(sf)
library(janitor)
library(lwgeom)

# Bigquery project
ucsb_project <-  "ucsb-gfw"

```

## Step 1: Isolating MPA boundaries

We downloaded the global dataset of protected areas (marine and terrestrial) from https://www.protectedplanet.net/ on November 13, 2019. 

As we are only interested in marine protected areas, we first need to filter this dataset to remove completely terrestrial protected areas (MARINE != 0). Due to the size of the global protected areas dataset, this was done using QGIS prior to importing into R. We then uploaded the resulting shapefile to mapshaper.org and simplified it to 0.5%. The resulting file is stored in `./data/WDPA_Nov2019_No_Terrestrial/`. 

```{r}
# Read in MPA data and convert to text format for upload to Bigquery
mpa_boundaries <- st_read(here::here("data", "WDPA_Nov2019_No_Terrestrial/"), layer = "WDPA_Nov2019_No_Terrestrial", stringsAsFactors=FALSE, as_tibble=TRUE) %>%
  clean_names(case = "snake") %>%
  dplyr::select(wdpa_pid, area_km = rep_m_area, no_take, area_no_take_km = no_tk_area, status, year = status_yr, loc_parent_iso3 = parent_iso, loc_iso3 = iso3) %>%
  st_make_valid() %>%
  mutate(mpa_boundary = st_as_text(geometry)) %>%
  st_set_geometry(NULL)

# Filter for only MPAs that include some portion as a no-take
no_take_mpa_boundaries <- mpa_boundaries %>%
  dplyr::filter(no_take %in% c("Part", "All"))

# Filter out ones larger than some area - may have to play around with what this area is for now... 
no_take_lsmpa_boundaries <- no_take_mpa_boundaries %>%
  dplyr::filter(area_km > 10000)

write_csv(no_take_lsmpa_boundaries, path = here::here("results", "LSMPAs", "no_take_lsmpa_boundaries.csv"))
```

For this analysis, we only extract MPAs where some portion has been designated as a "no-take" reserve, and we only consider reserves with a total size greater than 10,000 km2. 

```{r}
# Upload LSMPA boundaries to BQ
lsmpa_table <- bq_table(project = ucsb_project, table = "lsmpa_boundaries", dataset = "ocean_halos_v2") 

if(bq_table_exists(lsmpa_table) == F){
  
  lsmpa_table %>%
      bq_table_create()
}else{
  
  lsmpa_table %>%
    bq_table_delete() %>%
    bq_table_upload(values = no_take_lsmpa_boundaries)
}
```

```{r}
# Set grid resolution
resolution <- 0.1
lats <- seq(-90 + resolution / 2, 90 - resolution / 2, resolution)
```

```{r}
# Make EEZ boundaries and shape spatial
sql<-"#standardSQL
SELECT mrgid_eez,
sovereign_iso3,
territory_iso3,
ST_GEOGFROMTEXT(eez_boundary) eez_boundary
FROM `ucsb-gfw.ocean_halos.eez_200nm_boundaries`
LEFT JOIN
(SELECT mrgid_eez,sovereign_iso3,territory_iso3 FROM `ucsb-gfw.ocean_halos.eez_info_lookup`)
USING (mrgid_eez)"

# Upload EEZ boundaries to BQ
# bq_table(project = ucsb_project, table = "eez_200nm_boundaries_spatial", dataset = "ocean_halos") %>% 
#   bq_table_create()
bq_table(project = ucsb_project, table = "eez_200nm_boundaries_spatial", dataset = "ocean_halos") %>% 
  bq_table_delete()
bq_project_query(ucsb_project,sql, destination_table = bq_table(project = ucsb_project, table = "eez_200nm_boundaries_spatial", dataset = "ocean_halos"), use_legacy_sql = FALSE, allowLargeResults = TRUE)
```

```{r}
# Pull in gridded EEZ info
# sql<-"#standardSQL
#   WITH 
#   crossed_gridcode AS(
#   SELECT
#     CAST(SUBSTR(gridcode, 5, 7) AS FLOAT64) lon,
#     CAST(SUBSTR(gridcode, 17, 7) AS FLOAT64) lat,
#     CAST(mrgid_eez AS INT64) mrgid_eez
#   FROM
#     `world-fishing-827.pipe_static.spatial_measures_20181025`
#   LEFT JOIN
#     UNNEST(regions.eez) mrgid_eez
#   WHERE
#     elevation_m <= 0),
# 
# grouped_grids AS(
#   SELECT
#     FLOOR(lat / 0.025) * 0.025 + 0.0125 lat_bin,
#     FLOOR(lon / 0.025) * 0.025 + 0.0125 lon_bin,
#     mrgid_eez,
#     COUNT(*) count
#   FROM
#     crossed_gridcode
#   GROUP BY
#     lat_bin,
#     lon_bin,
#     mrgid_eez),
# 
# grouped_grids_2 AS(
#   SELECT
#     *,
#     FORMAT('lon:%+07.2f_lat:%+07.2f',
#       lon_bin,
#       lat_bin) AS gridcode_binned
#   FROM
#     grouped_grids),
# 
# group_grids_3 AS(
#   SELECT
#     lat_bin,
#     lon_bin,
#     gridcode_binned,
#     mrgid_eez,
#     count,
#     ROW_NUMBER() OVER (PARTITION BY gridcode_binned ORDER BY count DESC) AS eez_rank
#   FROM
#     grouped_grids_2 )
# SELECT
#   *
# FROM
#   group_grids_3
# WHERE lat_bin <90 AND lat_bin > -90 AND lon_bin < 180 AND lon_bin > -180"

# We're just going to pull this from Gavin's table because I don't have access to this.
sql<-"#standardSQL
SELECT
  *
FROM
  `ucsb-gfw.eez_analysis.grid_eez_locations`"


bq_table(project = ucsb_project, table = "grid_eez_locations", dataset = "ocean_halos") %>% 
  bq_table_create()
bq_table(project = ucsb_project, table = "grid_eez_locations", dataset = "ocean_halos") %>% 
  bq_table_delete()
bq_project_query(ucsb_project, sql, destination_table = bq_table(project = ucsb_project, table = "grid_eez_locations", dataset = "ocean_halos"), use_legacy_sql = FALSE, allowLargeResults = TRUE)
```

```{r}
# Condense this down to relevant areas
sql<-"#standardSQL
WITH
  multi_grids AS(
  SELECT
    gridcode_binned
  FROM
    `ucsb-gfw.ocean_halos.grid_eez_locations`
  WHERE
    eez_rank > 1
  GROUP BY
    gridcode_binned ),

eez_grids AS(
  SELECT
    lat_bin,
    lon_bin,
    ST_GeogPoint(lon_bin,
      lat_bin) grid_location,
    gridcode_binned,
    mrgid_eez
  FROM (
    SELECT
      *
    FROM
      `ucsb-gfw.ocean_halos.grid_eez_locations`
    WHERE
      NOT gridcode_binned IN (
      SELECT
        gridcode_binned
      FROM
        multi_grids)))
SELECT
  *
FROM
  eez_grids"

# bq_table(project = ucsb_project, table = "grid_eez_locations_condensed", dataset = "ocean_halos") %>% 
#   bq_table_create()
bq_table(project = ucsb_project,table = "grid_eez_locations_condensed",dataset = "ocean_halos") %>% 
  bq_table_delete()
bq_project_query(ucsb_project,sql, destination_table = bq_table(project = ucsb_project,table = "grid_eez_locations_condensed",dataset = "ocean_halos"),use_legacy_sql = FALSE, allowLargeResults = TRUE)
```

```{r}
# Get distances for points inside relevant EEZs
sql<-paste0("#standardSQL
WITH
  eez_grids AS(
  SELECT
    lat_bin,
    lon_bin,
    grid_location,
    gridcode_binned,
    mrgid_eez,
    sovereign_iso3,
    territory_iso3,
    eez_boundary
  FROM `ucsb-gfw.ocean_halos.grid_eez_locations_condensed`
  JOIN (
    SELECT
      *
    FROM
      `ucsb-gfw.ocean_halos.eez_200nm_boundaries_spatial`)
  USING
    (mrgid_eez))

SELECT
  lat_bin,
  lon_bin,
  gridcode_binned,
  mrgid_eez,
  sovereign_iso3 closest_eez_boundary_sovereign,
  territory_iso3 closest_eez_boundary_territory,
  TRUE inside_eez,
  ST_Distance(eez_boundary, grid_location) distance_to_eez_boundary_meters
FROM
  eez_grids
WHERE
 mrgid_eez IN (", SID_codes_string, ")")

# bq_table(project = ucsb_project,table = "master_table_inside_eez_SID",dataset = "ocean_halos") %>% 
#   bq_table_create()
bq_table(project = ucsb_project,table = "master_table_inside_eez_SID",dataset = "ocean_halos") %>% 
  bq_table_delete()
bq_project_query(ucsb_project, sql, destination_table = bq_table(project = ucsb_project, table = "master_table_inside_eez_SID",dataset = "ocean_halos"),use_legacy_sql = FALSE, allowLargeResults = TRUE)
```


```{r}
sql<-paste0("#standardSQL
  WITH high_seas_grids AS(
  SELECT
    lat_bin,
    lon_bin,
    grid_location,
    gridcode_binned
  FROM
    `ucsb-gfw.ocean_halos.grid_eez_locations_condensed`
  WHERE
    mrgid_eez IS NULL),

  eez_grids AS(
  SELECT
    *
  FROM
    high_seas_grids
  CROSS JOIN (
    SELECT
    eez_boundary,
    mrgid_eez,
    sovereign_iso3,
    territory_iso3
    FROM
      `ucsb-gfw.ocean_halos.eez_200nm_boundaries_spatial`)),

  all_info AS(
  SELECT
    lat_bin,
    lon_bin,
    gridcode_binned,
    mrgid_eez,
    sovereign_iso3 closest_eez_boundary_sovereign,
    territory_iso3 closest_eez_boundary_territory,
    FALSE inside_eez,
    ST_Distance(eez_boundary,
      grid_location) distance_to_eez_boundary_meters
  FROM
    eez_grids),

  ranked_info AS(
  SELECT
    lat_bin,
    lon_bin,
    gridcode_binned,
    mrgid_eez,
    closest_eez_boundary_sovereign,
    closest_eez_boundary_territory,
    inside_eez,
    distance_to_eez_boundary_meters,
    ROW_NUMBER() OVER (PARTITION BY gridcode_binned ORDER BY distance_to_eez_boundary_meters ASC) AS eez_rank
  FROM
    all_info)

SELECT
  lat_bin,
  lon_bin,
  gridcode_binned,
  mrgid_eez,
  closest_eez_boundary_sovereign,
  closest_eez_boundary_territory,
  inside_eez,
  distance_to_eez_boundary_meters
FROM
  ranked_info
WHERE
  eez_rank = 1
  AND mrgid_eez IN (", SID_codes_string, ")")

# bq_table(project = ucsb_project,table = "master_table_high_seas_SID",dataset = "ocean_halos") %>% 
#   bq_table_create()
bq_table(project = ucsb_project,table = "master_table_high_seas_SID",dataset = "ocean_halos") %>% 
  bq_table_delete()
bq_project_query(ucsb_project,sql, destination_table = bq_table(project = ucsb_project,table = "master_table_high_seas_SID",dataset = "ocean_halos"),use_legacy_sql = FALSE, allowLargeResults = TRUE)
```

```{r}
# Combine master tables into one for SIDs
# Merge inside and high seas tables
sql <- "
SELECT
*
FROM
`ucsb-gfw.ocean_halos.master_table_inside_eez_SID`
UNION ALL
(SELECT
*
FROM
`ucsb-gfw.ocean_halos.master_table_high_seas_SID`)"

bq_table(project = ucsb_project,table = "master_table_all_SID",dataset = "ocean_halos") %>% 
  bq_table_create()
bq_table(project = ucsb_project,table = "master_table_all_SID",dataset = "ocean_halos") %>% 
  bq_table_delete()
bq_project_query(ucsb_project,sql, destination_table = bq_table(project = ucsb_project,table = "master_table_all_SID",dataset = "ocean_halos"),use_legacy_sql = FALSE, allowLargeResults = TRUE)

```

```{r}
test_eez <- 8440
# Make test map of FP to see how this worked
test <- bq_project_query(ucsb_project, "SELECT * FROM `ucsb-gfw.ocean_halos.master_table_high_seas` WHERE mrgid_eez = 8440", allowLargeResults = TRUE) %>%
  bq_table_download(max_results = Inf)

test_boundary <- st_read(here::here("data", "World_EEZ_v10_20180221/"), layer="eez_boundaries_v10", stringsAsFactors=FALSE, as_tibble=TRUE) %>%
    filter(Line_type == "200 NM") %>%
  dplyr::select(mrgid_eez = MRGID_EEZ1, territory = Territory1, sovereign = Sovereign1) %>%
  filter(mrgid_eez == test_eez)

world_land <-
  st_as_sf(rworldmap::countriesLow) ## data(countriesLow) is from the rworldmap package

test %>%
  mutate(distance_to_eez_boundary_nm = distance_to_eez_boundary_meters / 1852) %>%
  filter(distance_to_eez_boundary_nm < 200) %>%
  #filter(mrgid_eez == test_eez)  %>%
  ggplot()+
  geom_tile(aes(x = lon_bin, y=lat_bin, fill = distance_to_eez_boundary_nm)) +
  coord_map(projection = "mollweide") +
  geom_sf(data=world_land, color = NA, fill="grey30") +
  #geom_sf(data=test_boundary,color="red") +
  viridis::scale_fill_viridis("[nm]") +
  xlab("") +
  ylab("") +
  theme(plot.title = element_text(color="black",hjust=0,vjust=1, size=15),
        panel.background = element_rect(fill ="black"),
        legend.text = element_text(color = "black", size = 15),
        legend.title = element_text(color = "black", size = 15),
        legend.title.align = 1,
        legend.background = element_rect(fill="white"),
        legend.direction="vertical",
        legend.justification = "center",
        axis.text = element_text(color = "black", size = rel(1)),
        panel.grid.major = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  ggtitle("Distance to all 200nm EEZ boundaries, capped at 200nm\n(full dataset has no distance cap)")

```

```{r}
# Pull in fishing data from 2016, 2017, 2018

# sql<-"#standardSQL
# WITH
#  vessel_info AS(
#   SELECT
#     ssvid,
#     year,
#     best.best_vessel_class gear,
#     best.best_flag flag
#   FROM
#     `world-fishing-827.gfw_research.vi_ssvid_byyear`
#   WHERE
#    on_fishing_list_best
#    AND activity.offsetting IS FALSE),
#      
# base AS(
#   SELECT
#     ssvid,
#     EXTRACT(year FROM timestamp) year,
#     FLOOR(lat / 0.025) * 0.025 + 0.0125 lat_bin,
#     FLOOR(lon / 0.025) * 0.025 + 0.0125 lon_bin,
#     IF(distance_from_port_m >= 1000 AND speed >= 1 AND nnet_score2 = 1, hours, NULL) fishing_hours
#   FROM
#     `world-fishing-827.gfw_research.pipe_production_b_fishing`
#   WHERE
#     _PARTITIONTIME BETWEEN TIMESTAMP('2016-01-01')
#     AND TIMESTAMP('2018-12-31')
#     AND seg_id IN (
#         SELECT
#         seg_id
#         FROM
#          `world-fishing-827.gfw_research.pipe_production_b_segs`
#         WHERE
#         good_seg)),
#   
# base_binned AS(
#   SELECT
#     lat_bin,
#     lon_bin,
#     year,
#     ssvid,
#     SUM(fishing_hours) fishing_hours
#   FROM
#     base
#   GROUP BY
#     lat_bin,
#     lon_bin,
#     year,
#     ssvid),
#     
# base_joined AS(
#   SELECT
#     *
#   FROM
#     base_binned
#   JOIN
#     vessel_info
#   USING
#     (ssvid,
#      year))
#      
# SELECT
#   lat_bin,
#   lon_bin,
#   year,
#   flag,
#   gear,
#   SUM(fishing_hours) fishing_hours
# FROM
#   base_joined
# GROUP BY
#   lat_bin,
#   lon_bin,
#   year,
#   flag,
#   gear"

# bq_table(project = ucsb_project,table = "fishing_data_2016_2018",dataset = "eez_analysis") %>% 
#   bq_table_delete()
# bq_project_query(ucsb_project,sql, destination_table = bq_table(project = ucsb_project,table = "fishing_data_2016_2018",dataset = "eez_analysis"),use_legacy_sql = FALSE, allowLargeResults = TRUE)

# ^ Did this chunk in BQ directly because it's easier and faster... 
```

```{r}
# Combine high seas bins with fishing information
sql<-"#standardSQL
 WITH distance_info AS(
  SELECT
    lat_bin,
    lon_bin,
    gridcode_binned,
    mrgid_eez,
    closest_eez_boundary_sovereign as eez_boundary_sovereign,
    closest_eez_boundary_territory as eez_boundary_territory,
    inside_eez,
    distance_to_eez_boundary_meters
  FROM
    `ucsb-gfw.ocean_halos.master_table_all_SID`
  WHERE
    NOT closest_eez_boundary_sovereign IS NULL),

fishing_info AS(
  SELECT
    lat_bin,
    lon_bin,
    year,
    SUM(fishing_hours) fishing_hours
  FROM
    `ucsb-gfw.ocean_halos.fishing_data_2016_2018`
  WHERE NOT fishing_hours IS NULL
  GROUP BY
    lat_bin,
    lon_bin,
    year),

joined_table AS(
SELECT
  *
FROM
  distance_info
JOIN
  fishing_info
USING(lat_bin,
    lon_bin))

SELECT
  lat_bin,
  lon_bin,
  year,
  mrgid_eez,
  eez_boundary_sovereign,
  eez_boundary_territory,
  inside_eez,
  distance_to_eez_boundary_meters,
  SUM(fishing_hours) fishing_hours
FROM
  joined_table
GROUP BY
  lat_bin,
  lon_bin,
  year,
  mrgid_eez,
  eez_boundary_sovereign,
  eez_boundary_territory,
  inside_eez,
  distance_to_eez_boundary_meters
ORDER BY
  year,
  lat_bin,
  lon_bin,
  mrgid_eez,
  eez_boundary_sovereign,
  eez_boundary_territory,
  distance_to_eez_boundary_meters
"
bq_table(project = ucsb_project,table = "full_dataset_SID",dataset = "ocean_halos") %>%
  bq_table_create()
bq_table(project = ucsb_project,table = "full_dataset_SID",dataset = "ocean_halos") %>% 
  bq_table_delete()
bq_project_query(ucsb_project,sql, destination_table = bq_table(project = ucsb_project,table = "full_dataset_SID",dataset = "ocean_halos"),use_legacy_sql = FALSE, allowLargeResults = TRUE)
```

```{r}
# Minimum distances
sql<-"SELECT
closest_eez_boundary_territory,
closest_eez_boundary_sovereign,
inside_eez ,
MIN( distance_to_eez_boundary_meters ) min_distance
FROM
  `ucsb-gfw.ocean_halos.master_table_all_SID`
  WHERE NOT closest_eez_boundary_territory IS NULL
GROUP BY
closest_eez_boundary_territory,
closest_eez_boundary_sovereign ,
inside_eez
ORDER BY
closest_eez_boundary_territory,
inside_eez,
min_distance"

min_distances_by_eez <- bq_project_query(ucsb_project, sql) %>%
  bq_table_download(max_results = Inf)

write_csv(min_distances_by_eez, path = here::here("data", "min_distances_by_eez.csv"))
```

```{r}
# Find grid cells that are overlapping the line, split between inside and outside
sql<-"#standardSQL
  CREATE TEMP FUNCTION RADIANS(x FLOAT64) AS ( ACOS(-1) * x / 180 ); CREATE TEMP FUNCTION RADIANS_TO_KM(x FLOAT64) AS ( 111.045 * 180 * x / ACOS(-1) ); CREATE TEMP FUNCTION HAVERSINE(lat1 FLOAT64,
    long1 FLOAT64,
    lat2 FLOAT64,
    long2 FLOAT64) AS (( RADIANS_TO_KM( ACOS(COS(RADIANS(lat1)) * COS(RADIANS(lat2)) * COS(RADIANS(long1) - RADIANS(long2)) + SIN(RADIANS(lat1)) * SIN(RADIANS(lat2)))) ) * 1000); 
WITH expanded AS(
  SELECT
    *,
    HAVERSINE(lat_bin,
      lon_bin-0.05,
      lat_bin,
      lon_bin+0.05) x_dist_m,
    HAVERSINE(lat_bin-0.05,
      lon_bin,
      lat_bin+0.05,
      lon_bin) y_dist_m
  FROM
    `ucsb-gfw.ocean_halos.full_dataset_SID`),
  master AS(
  SELECT
    * EXCEPT(fishing_hours,
      inside_eez,
      x_dist_m,
      y_dist_m),
    fishing_hours fishing_hours,
    CAST(inside_eez AS BOOL) inside_eez,
    FALSE split_grid
  FROM
    expanded
  WHERE
    distance_to_eez_boundary_meters >= (GREATEST(x_dist_m,y_dist_m)/2)),
  
filtered AS(
  SELECT
    * EXCEPT(fishing_hours,
    distance_to_eez_boundary_meters,
      inside_eez,
      x_dist_m,
      y_dist_m),
      (2 * distance_to_eez_boundary_meters + GREATEST(x_dist_m,y_dist_m)) / 4 distance_to_eez_boundary_meters,
    fishing_hours * distance_to_eez_boundary_meters / (GREATEST(x_dist_m,y_dist_m)/2) fishing_hours,
    CAST(inside_eez AS BOOL) inside_eez,
    TRUE split_grid
  FROM
    expanded
  WHERE
    distance_to_eez_boundary_meters < (GREATEST(x_dist_m,y_dist_m)/2)),
  filtered_flipped AS(
  SELECT
    * EXCEPT(fishing_hours,
    distance_to_eez_boundary_meters,
      inside_eez,
      x_dist_m,
      y_dist_m),
      (GREATEST(x_dist_m,y_dist_m) - 2 * distance_to_eez_boundary_meters) / 4 distance_to_eez_boundary_meters,
    fishing_hours * (1 - distance_to_eez_boundary_meters / (GREATEST(x_dist_m,y_dist_m)/2)) fishing_hours,
    NOT(CAST(inside_eez AS BOOL)) inside_eez,
    TRUE split_grid
  FROM
    expanded
  WHERE
    distance_to_eez_boundary_meters < (GREATEST(x_dist_m,y_dist_m)/2)),

joined AS(
  SELECT
    *
  FROM
    filtered UNION ALL (
    SELECT
      *
    FROM
      master) UNION ALL (
    SELECT
      *
    FROM
      filtered_flipped))
SELECT
  *
FROM
  joined"

bq_table(project = ucsb_project,table = "final_eez_dataset_SID",dataset = "ocean_halos") %>% 
  bq_table_create()
bq_table(project = ucsb_project,table = "final_eez_dataset_SID",dataset = "ocean_halos") %>% 
  bq_table_delete()
bq_project_query(ucsb_project,sql, destination_table = bq_table(project = ucsb_project,table = "final_eez_dataset_SID", dataset = "ocean_halos"),use_legacy_sql = FALSE, allowLargeResults = TRUE)
```


```{r}
# final_eez_dataset <- bq_project_query(ucsb_project, "SELECT * FROM `ucsb-gfw.ocean_halos.full_dataset`") %>%
#   bq_table_download(max_results = Inf)
# 
# write_csv(final_eez_dataset, path = here::here("data", "final_eez_dataset.csv"))
```

